Sparse spectral and $p$-finite element methods on disk-slices and rhombi I: zero boundary conditions

### 2019-02-22 papers/halfdisk/halfdisk.tex
Further edits to paper

Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^T \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to. Note that due to orthogonality this is a diagonal matrix. We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
\begin{align}
\bigWii(x,y) &:= \begin{bmatrix}
		\bigWii_0(x,y) \\
		\bigWii_1(x,y) \\
		\bigWii_2(x,y) \\
		\vdots \\
		\bigWii_p(x,y)
	\end{bmatrix}, \\
\bigWii_n(x,y) &:= \begin{bmatrix}
		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
		\vdots \\
		\Wii(a,y) \: \Pii_{n,n}(x,y)
	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
\end{align}
and rewrite (63) in matrix form:
\begin{align}
	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
	&= \int_\element \begin{bmatrix}
					\partial_x v_h \\
					\partial_y v_h
				\end{bmatrix}^T 
				\begin{bmatrix}
					\partial_x u_h \\
					\partial_y u_h
				\end{bmatrix}
				\\
	&= \int_\element \begin{bmatrix}
					\bigPoo^T \Wii_x \mathbf{v} \\
					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
				\end{bmatrix}^T 
				\begin{bmatrix}
					\bigPoo^T \Wii_x \mathbf{u} \\
					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
				\end{bmatrix}
				\\
	&= \int_\element \Big( \mathbf{v}^T {\Wii_x}^T \bigPoo \bigPoo^T \Wii_x \mathbf{u} \nonumber \\
					& \quad \quad \quad + \mathbf{v}^T ({T^{(1,0)\to(0,0)} \Wii_y})^T \bigPoo \bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
	&= \mathbf{v}^T \: \Big( {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
\end{align}
where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
\begin{align}
	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
\end{align}
where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis. We can also not that naturally $\langle \bigPii, {\bigPii}^T \rangle_{\Wii}$ is just a diagonal matrix.

Since (63) is equivalent to stating that
\begin{align}
	L(\Wii \Pii_{n,k}) = a(u_h,\Wii \Pii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
\end{align}
(i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
\begin{align}
A\mathbf{u} = \tilde{\mathbf{f}}.
\end{align}
where the (element) stiffness matrix $A$ is defined by 
\begin{align}
A = {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
\end{align}
and the load vector $\tilde{\mathbf{f}}$ is given by 
\begin{align}
\tilde{\mathbf{f}} = \Lambda^{(1,1)} \: \mathbf{f}.
\end{align}

@@

\section{Disk Slices}

The work in this paper on the half disk can be easily transferred to the domain of a disk slice by which we mean
\begin{align}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align}
with $\alpha, \beta \in [0,1]$, $\rho(x) = \sqrt{1-x^2}$, and  $-\gamma = \delta  = $

>>>>

Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^T \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to. Note that due to orthogonality this is a diagonal matrix. We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
\begin{align}
\bigWii(x,y) &:= \begin{bmatrix}
		\bigWii_0(x,y) \\
		\bigWii_1(x,y) \\
		\bigWii_2(x,y) \\
		\vdots \\
		\bigWii_p(x,y)
	\end{bmatrix}, \\
\bigWii_n(x,y) &:= \begin{bmatrix}
		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
		\vdots \\
		\Wii(a,y) \: \Pii_{n,n}(x,y)
	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
\end{align}
and rewrite (73) in matrix form:
\begin{align}
	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
	&= \int_\element \begin{bmatrix}
					\partial_x v_h \\
					\partial_y v_h
				\end{bmatrix}^T 
				\begin{bmatrix}
					\partial_x u_h \\
					\partial_y u_h
				\end{bmatrix}
				\\
	&= \int_\element \begin{bmatrix}
					\bigPoo^T \Wii_x \mathbf{v} \\
					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
				\end{bmatrix}^T 
				\begin{bmatrix}
					\bigPoo^T \Wii_x \mathbf{u} \\
					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
				\end{bmatrix}
				\\
	&= \int_\element \Big( \mathbf{v}^T {\Wii_x}^T \bigPoo \bigPoo^T \Wii_x \mathbf{u} \nonumber \\
					& \quad \quad \quad + \mathbf{v}^T ({T^{(1,0)\to(0,0)} \Wii_y})^T \bigPoo \bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
	&= \mathbf{v}^T \: \Big( {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
\end{align}
where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
\begin{align}
	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
	&= \mathbf{v}^T \: \langle \bigPii, {\bigPii}^T \rangle_{\Wii} \: d\mathbf{x} \\
	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
\end{align}
where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis.

Since (73) is equivalent to stating that
\begin{align}
	L(\Wii \Pii_{n,k}) = a(u_h,\Wii \Pii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
\end{align}
(i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
\begin{align}
A\mathbf{u} = \tilde{\mathbf{f}}.
\end{align}
where the (element) stiffness matrix $A$ is defined by 
\begin{align}
A = {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
\end{align}
and the load vector $\tilde{\mathbf{f}}$ is given by 
\begin{align}
\tilde{\mathbf{f}} = \Lambda^{(1,1)} \: \mathbf{f}.
\end{align}

@@

\section{Disk slices}

The work in this paper on the half disk can be easily transferred to the domain of a disk slice by which we mean
\begin{align}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align}
with 
\begin{align}
\begin{cases}
-1 \le \alpha < &\beta \le 1 \\
(\gamma, \delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w^{(a,b,c)}_1(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c} = (\beta - x)^a \: (x - \alpha)^{b} \: (1-x^2)^{c} \\
w^{(b)}_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b.
\end{cases}
\end{align}
The weight $w^{(b)}_2(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(b, b)}\}$. $w^{(a,b,c)}_1(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a, b,c)}\}\). Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(c, a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal with respect to the weight
\begin{align}
	W^{(a,b,c)}(x,y) &:= w^{(c,a,b)}_1(x) w^{(b)}_2(\frac{y}{\rho(x)}) \nonumber \\
	&= (\beta - x)^c \: (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
	&= (\beta - x)^c \: (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
\end{align}

Note that by setting $c = 0$, $\alpha = 0$ and $\beta = 1$ we obtain the 2D orthogonal polynomials on the half disk.

\section{Hemisphere}

We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
